---
title: 'Data Transformation: Iteration 2'
author: "Carson Slater"
date: '2022-10-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r, include=FALSE}
# Loading packages
library(tidymodels)
library(stringr)
library(janitor)
library(glmnet)
library(lubridate)
library(knitr)
library(mosaic)
library(doParallel)
```
```{r}
# Insert pathname below to load any data
path <- "/Users/carson/Documents/Wheaton Senior Year/Fall 2022/Mentored Research/new_baskets_adhie.csv"

df <- read.csv(path)
```

# Data Transformation Steps for the `Prophet` Modeling Process
### Preparing the Data with Robust Code (For Awantunai Data)
```{r}
# Finding Percentage of the Missing Data for each column.
colMeans(is.na(df))*100
# There is a very small proportion of missing data in these data 

df <- df %>% mutate(id = as.factor(id), 
                        order_id = as.factor(order_id), 
                        merchant_id = as.factor(merchant_id), 
                        sku_id = as.factor(sku_id), 
                        top_cat_id = as.factor(top_cat_id), 
                        sub_cat_id = as.factor(sub_cat_id))

# Cleaning data so that R can read the time stamp
df$placed_at = substr(df$placed_at, 1, nchar(df$placed_at)-4)

# Changing the placed_by into a POSIXct variable type
df$placed_at = as.POSIXct(df$placed_at)

# Finding NA's
df %>% filter(is.na(df$top_cat_id))

# Removing the 11 NA's
df <- df %>% filter(!is.na(df$top_cat_id))


# Creating more columns with more date-specific information
df <- df %>% mutate(year = format(df$placed_at, "%Y"),
                        month = format(df$placed_at, "%m"),
                        day = format(df$placed_at, "%d"),
                        hour = format(df$placed_at, "%H"),
                        minute = format(df$placed_at, "%M"),
                        second = format(df$placed_at, "%S"),
                        yday = yday(df$placed_at),
                        wday = wday(df$placed_at),
                        yweek = week(df$placed_at))

# Creating factor variables for dates and times
df <- df %>% mutate(year = as.factor(year),
                        month = as.factor(month),
                        day = as.factor(day),
                        hour = as.factor(hour),
                        minute = as.factor(minute),
                        second = as.factor(second),
                        yday = as.factor(yday),
                        wday = as.factor(wday),
                        yweek = as.factor(yweek))

# Looking for duplicates orders
dupes <- get_dupes(df, order_id, placed_at, merchant_id, sku_id)

# Removing the duplicate observations
df <- df %>% distinct(order_id, placed_at, merchant_id, sku_id, .keep_all = TRUE)
```

### Exploratory Data Analysis
From my first iteration I learned that there is no 'one size fits all' forecasting model. I want to see if I can create a data frame that will contain all of the summary statistics for each SKU in a data set, containing range, mean order volume, median order volume, etc. 

```{r}
# Creating a data frame with the summary statistics for each SKU
agg_vol <- df %>%
  group_by(sku_id) %>% 
  summarize(tot_qty = sum(qty), ) %>% 
  arrange(desc(tot_qty))
```



