---
title: 'Data Transformation: Iteration 2'
author: "Carson Slater"
date: '2022-10-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r, include=FALSE}
# Loading packages
library(tidymodels)
library(stringr)
library(janitor)
library(glmnet)
library(lubridate)
library(knitr)
library(mosaic)
library(doParallel)
```
```{r, eval=FALSE}
# Insert pathname below to load any data
path <- "/Users/carson/Documents/Wheaton Senior Year/Fall 2022/Mentored Research/new_baskets_adhie.csv"

df <- read.csv(path)
```

# Data Transformation Steps for the `Prophet` Modeling Process
### Preparing the Data with Robust Code (For Awantunai Data)
```{r, eval=FALSE}
# Finding Percentage of the Missing Data for each column.
colMeans(is.na(df))*100
# There is a very small proportion of missing data in these data 

df <- df %>% mutate(id = as.factor(id), 
                        order_id = as.factor(order_id), 
                        merchant_id = as.factor(merchant_id), 
                        sku_id = as.factor(sku_id), 
                        top_cat_id = as.factor(top_cat_id), 
                        sub_cat_id = as.factor(sub_cat_id))

# Cleaning data so that R can read the time stamp
df$placed_at = substr(df$placed_at, 1, nchar(df$placed_at)-4)

# Changing the placed_by into a POSIXct variable type
df$placed_at = as.Date(df$placed_at)

# Finding NA's
df %>% filter(is.na(df$top_cat_id))

# Removing the 11 NA's
df <- df %>% filter(!is.na(df$top_cat_id))


# Creating more columns with more date-specific information
df <- df %>% mutate(year = format(df$placed_at, "%Y"),
                        month = format(df$placed_at, "%m"),
                        day = format(df$placed_at, "%d"),
                        hour = format(df$placed_at, "%H"),
                        minute = format(df$placed_at, "%M"),
                        second = format(df$placed_at, "%S"),
                        yday = yday(df$placed_at),
                        wday = wday(df$placed_at),
                        yweek = week(df$placed_at))

# Creating factor variables for dates and times
df <- df %>% mutate(year = as.factor(year),
                        month = as.factor(month),
                        day = as.factor(day),
                        hour = as.factor(hour),
                        minute = as.factor(minute),
                        second = as.factor(second),
                        yday = as.factor(yday),
                        wday = as.factor(wday),
                        yweek = as.factor(yweek))

# Looking for duplicates orders
dupes <- get_dupes(df, order_id, placed_at, merchant_id, sku_id)

# Removing the duplicate observations
df <- df %>% distinct(order_id, placed_at, merchant_id, sku_id, .keep_all = TRUE)
```

### Exploratory Data Analysis
From my first iteration I learned that there is no 'one size fits all' forecasting model. I want to see if I can create a data frame that will contain all of the summary statistics for each SKU in a data set, containing range, mean order volume, median order volume, etc. 

```{r, eval=FALSE}
# Creating a data frame with the summary statistics for each SKU
sku_cnt <- df %>% count(sku_id) %>% filter(n >= 20)
  
sum_stats <- df %>% filter(df$sku_id %in% sku_cnt$sku_id) %>% 
  mutate(n = count(sku_id)) %>% 
  group_by(sku_id) %>%
  summarize(tot_vol = sum(qty),
            avg_qty = mean(qty),
            variance = var(qty),
            quartile1 = quantile(qty, 0.25),
            quartile3 = quantile(qty, 0.75),
            med_qty = median(qty),
            min = min(qty), 
            max = max(qty),
            range = max - min) %>% 
  distinct %>% 
  arrange(desc(tot_vol))
```

Here I think I have computed some summary statistics for these data. I want to dive deeper into these data to see if there are any noteworthy finds for each SKU.

```{r, eval=FALSE}
# total volume
sum_stats %>% arrange(desc(tot_vol)) %>% head(15) %>% kable()

sku_cnt %>% arrange(desc(n)) %>% head(5) %>% kable()

sum_stats %>% filter(sku_id == 1488)
sum_stats %>% filter(sku_id == 859)
sum_stats %>% filter(sku_id == 1497)
sum_stats %>% filter(sku_id == 866)
sum_stats %>% filter(sku_id == 976)
# most of these high frequency SKU's have very low variability in order quantites

sum_stats %>% arrange(desc(med_qty), desc(variance))
```

I need to transform the data and aggregate by day each order volume.

```{r, eval=FALSE}
by_day <- df %>% select(placed_at, qty, sku_id) %>% 
  rename(ds = placed_at, y = qty) %>%
  mutate(ds = as.Date(ds)) %>% 
  group_by(ds, sku_id) %>% 
  mutate(y = sum(y)) %>% 
  unique()

by_day_stats <- by_day %>% 
  mutate(n = count(sku_id)) %>% 
  group_by(sku_id) %>%
  summarize(tot_vol = sum(y),
            avg_qty = mean(y),
            variance = var(y),
            quartile1 = quantile(y, 0.25),
            quartile3 = quantile(y, 0.75),
            med_qty = median(y),
            min = min(y), 
            max = max(y),
            range = max - min) %>% 
  distinct %>% 
  arrange(desc(tot_vol))
```

```{r, eval=FALSE}
save.image("Data_Iteration2.Rdata")
```
```{r, eval=FALSE}
load("Data_Iteration2.Rdata")
```

#### HALF-BAKED IDEA
Detect out liars by computing the 90-10 interval. Any outside of that can be ignored.

*I need to revisit this idea and make modifications*

