---
title: "Forecasting SKU Demand with Prophet"
author: "Carson Slater"
date: '2022-09-21'
output: html_document
header-includes:
   - \usepackage{amsmath}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r, include=FALSE}
# Loading packages
library(tidymodels)
library(tinytex)
library(stringr)
library(janitor)
library(glmnet)
library(lubridate)
library(knitr)
library(mosaic)
library(doParallel)
```

```{r, include=FALSE}
# For parallel processing
all_cores <- parallel::detectCores(logical = FALSE)
cl <- makePSOCKcluster(all_cores)
registerDoParallel(cl)
```

### Loading the Data
```{r}
load("SKU50.Rdata")
```

```{r}
# Splitting Each Data Frame
full_split <- full_50 %>% 
  group_by(sku_id) %>% 
  group_split()
```

### Making a Plot
```{r}
# For visualization purposes
library(RColorBrewer)
n <- 50
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]

col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, 
                           rownames(qual_col_pals)))

# The color schema I am using
pie(rep(1,n), col=sample(col_vector, n))
```

```{r}
qty_grp %>% ggplot(aes(y = sku_id, x = diff, fill = sku_id)) +
  geom_col(color = "black", show.legend = FALSE) +
  labs(title = "Desparity Between Mean and Median Order Volumes for the Top 50 SKU's",
       x = "SKU Identification Number",
       y = "Difference Between Mean and Median") +
  theme_minimal() +
  scale_fill_manual(values = col_vector)

qty_grp %>% ggplot(aes(x = avg_qty, y = med_qty)) +
  geom_point(color = "orange", show.legend = FALSE) +
  labs(title = "Desparity Between Mean and Median Order Volumes for the Top 50 SKU's",
       x = "Mean Order Volume",
       y = "Median Order Volume") +
  theme_minimal()
```

```{r, eval=FALSE}
library(prophet)
```

### Building Models
`Prophet` uses an additive regression model with four main components:

- A piecewise linear or logistic growth curve trend. Prophet automatically detects changes in trends by selecting changepoints from the data.
- A yearly seasonal component modeled using Fourier series.
- A weekly seasonal component using dummy variables.
- A user-provided list of important holidays.

The model looks like this:
\[
y(t) = g(t) + s(t) + h(t) + e(t)
\]

where,

- $g(t)$ refers to trend (changes over a long period of time)
- $s(t)$ refers to seasonality (periodic or short term changes)
- $h(t)$ refers to effects of holidays to the forecast
- $e(t)$ refers to the unconditional changes that is specific to a business or a person or a circumstance. It is also called the error term.
- $y(t)$ is the forecast.

Although prophet can fit a logistic growth model, its default and what we will be using for our purposes, is a piece-wise linear model,

\[   
y = \left\{
\begin{array}{ll}
     \beta_0 + \beta_1x, \hspace{2.5cm} x \leq c \\
     \beta_0 - \beta_2c + (\beta_1 + \beta_2)x, \hspace{4mm}x > c
\end{array} 
\right. 
\]

We can fit forecasting models for different types of SKU's within the top 50.

### Code for Analyzing an SKU Forecast
```{r}
# Preprocessing
df <-  full_split[[4]] # Choose any number in 1:50 to forecast a relevant SKU

df <- df %>% 
  select(c(placed_at, qty)) %>% 
  rename(ds = placed_at, y = qty) %>% 
  group_by(ds) %>% 
  summarise(y = sum(y), .groups = 'drop')

# Creating an evaluation data frame
eval <- df %>% filter(ds >= max(df$ds) - weeks(8)) %>% 
  complete(ds = seq.Date(min(ds), max(ds), by="day")) %>% 
  mutate(y = ifelse(is.na(y), 0, y))
  
eval <- eval %>% slice(1:(n() - 1))

orig_df <- df

df <- df %>% filter(ds < max(df$ds) - weeks(8))
```

```{r}
mod <- prophet(df, yearly.seasonality = TRUE, fit = FALSE, interval.width = 0.7) %>% 
  add_country_holidays(country_name = 'ID')

fit <- fit.prophet(mod, df)

# 6 Months
future <- make_future_dataframe(fit, periods = 8*7)

# Make a forecast
fcst <- predict(fit, future)

##########################################################
# Since we cannot have negaitve predictive quantities, 
# I invert the upper confidence interval and set the 
# negative yhats to zero
##########################################################

fcst <- fcst %>% mutate(diff = yhat_upper - yhat)

fcst <- fcst %>%
  mutate(yhat_upper = ifelse(yhat_upper >= 0, yhat_upper, diff)) %>% 
  mutate(yhat = ifelse(yhat >= 0, yhat, 0), 
         yhat_lower = ifelse(yhat_lower >= 0, yhat_lower, 0))

# Final plot of forecast
plot(fit, fcst) + 
  add_changepoints_to_plot(m = fit, cp_color = "orange") +
  ylim(0, 1750) +
  labs(title = "Fitted Model and Forecast of an SKU",
       x = "Date",
       y = "Volume Purchased") +
  theme_minimal()
```

```{r}
# Model Evaluation
pred <- fcst %>% filter(ds >= max(orig_df$ds) - weeks(8)) %>% 
  select(ds, yhat_lower, yhat, yhat_upper)

err_df <- tibble(pred$ds, eval$y, pred$yhat)

colnames(err_df) <- c("ds", "y", "yhat")

err_df <- err_df %>% mutate(resid = y - yhat)

#library(Metrics)
# Get RMSE
Metrics::rmse(err_df$y, err_df$yhat)

# Get SMAPE
Metrics::smape(err_df$y, err_df$yhat) # can be infinity if predicted\
# values in denominator are close to 1

err_df %>% 
  ggplot(aes(x = resid)) +
  geom_histogram(binwidth = 50, 
                 fill = "orange", 
                 color = "grey") +
  #xlim(-650,100) +
  labs(title = "Residual Plot",
       x = "Residual Value",
       y = "Frequency",
       caption = "SKU #200") +
  theme_minimal()
```
