---
title: "Keras Playground"
author: "Carson Slater"
date: '2022-10-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

### Introduction
Some of this was taken from [this article](https://www.kaggle.com/code/davidchilders/time-series-prediction-in-r-keras/notebook) from Kaggle.

I created this .Rmd file to play around with `Keras`, because every time I tried to import `Keras` in my jupyter notebook, my kernel crashed and I was unsure how to fix it without having to search the depths of my hard drive to solve the probable path issues.

```{r}
library(tidyverse)
library(keras)
library(knitr)
library(foreign) #Read files in Stata .dta format
library(reticulate) #Call Python from R
library(tensorflow) #Neural Network Backend for Keras
library(plyr) #Data manipulation
library(dplyr) # Data Manipulation
library(caret) #Machine Learning Tools for Training and Validation
```

### Importing the Data
This data was taken from the St. Louis branch of the federal reserve, on [this website](https://research.stlouisfed.org/econ/mccracken/fred-databases/).

```{r}
#Series of 729 observations of 129 macroeconomic indicators
url <- "https://files.stlouisfed.org/files/htdocs/fred-md/monthly/current.csv"
currentMD<-read.csv(url)
#Drop data column, so result contains only data
fmd<-select(currentMD,-one_of("sasdate"))
```

```{r}
#Convert to matrix (needed for replace_na, and later, for Keras)
mseries<-as.matrix(fmd)

#Extract indicators of missingness
maug<-1*data.frame(is.na(mseries)) #Multiplying by 1 converts logical to numeric

#Replace missing values with 0s, still needed because model can't handle NAs 
mseries<-data.frame(replace_na(mseries,0))

#Make feature matrix containing original features and missingness indicators for each feature
#Convert to arrays because Keras needs that
augseries<-as.matrix(bind_cols(mseries,maug))
```
```{r}
head(augseries)
```

```{r}
train_data<-augseries[1:521,]
mean <- apply(train_data, 2, mean)
std <- apply(train_data, 2, sd)
std<-case_when(std==0 ~ 1,
         std !=0 ~ std) #If data series is constant, recode standard deviation to 1, to avoid dividing by 0
data <- scale(augseries, center = mean, scale = std)

# Exactly copied generator function from Chollet and Allaire (p.195), 
# but setting 1 month steps and smaller batches for smaller data
# Also, allow which column of data is the one to be predicted to be an input
generator <- function(data, lookback, delay, min_index, max_index,
                      shuffle = FALSE, batch_size = 20, step = 1, predseries) {
  if (is.null(max_index)) max_index <- nrow(data) - delay - 1
  i <- min_index + lookback
  function() {
    if (shuffle) {
      rows <- sample(c((min_index+lookback):max_index), size = batch_size)
    } else {
      if (i + batch_size >= max_index)
        i <<- min_index + lookback
      rows <- c(i:min(i+batch_size, max_index))
      i <<- i + length(rows)
}
    samples <- array(0, dim = c(length(rows),
                                lookback / step,
                                dim(data)[[-1]]))
    targets <- array(0, dim = c(length(rows)))
    for (j in 1:length(rows)) {
      indices <- seq(rows[[j]] - lookback, rows[[j]],
                     length.out = dim(samples)[[2]])
      samples[j,,] <- data[indices,]
      targets[[j]] <- data[rows[[j]] + delay,predseries]
    }
    list(samples, targets)
  }
}
```

```{r}
head(data) %>% kable()
```

```{r}
lookback<- 48 #4 years of past data used in each current prediction
step <- 1 #Do not skip observations (could set to higher value if you had extremely high frquency data, like for stock market data)
delay<- 1 #Predict 1 month ahead
batch_size <- 10 #draw 20 samples at a time
predser <- 6 #Target is the 6th series in the list, Industrial Production

#Training set looks at first 521 months
train_gen <- generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 1,
  max_index = 521,
  shuffle = TRUE,
  step = step,
  batch_size = batch_size,
  predseries = predser  
)
#Validation set looks at next 119 months
val_gen = generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 522,
  max_index = 640,
  step = step,
  batch_size = batch_size,
  predseries = predser  
)
#Test set looks at remaining months (641-729)
test_gen <- generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 641,
  max_index = NULL,
  step = step,
  batch_size = batch_size,
  predseries = predser    
)

val_steps <- (640 - 521 - lookback) / batch_size
test_steps <- (nrow(data) - 641 - lookback) / batch_size
```

```{r}
set.seed(78910)
```

I cannot seem to get past this chunk, as each time I run it, the R session aborts.

```{r, eval=FALSE}
densemodel <- keras_model_sequential() %>%
    layer_flatten(input_shape=c(lookback/step,dim(data)[-1])) %>%
    layer_dense(units=32,activation="relu") %>%
    layer_dense(units=1)


densemodel %>% compile(
    optimizer = "rmsprop",
    loss="mse"
)

history <- densemodel %>% fit_generator(
  train_gen,
  steps_per_epoch = 47,
  epochs = 20,
  validation_data = val_gen,
  validation_steps = val_steps
)

densemodel %>% save_model_hdf5("densetimeseries.h5")
```

