---
title: "Exponential Smoothing"
author: "Carson Slater"
date: '2022-09-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

### Packages Considered
```{r}
library(tidymodels)
library(forecast)
library(lubridate)
library(zoo)
library(fpp2)
```

### Loading the Data
```{r}
load("SKU50.Rdata")
```

### Subsetting the Data
```{r}
# Obtaining a subset of the data for a particular good
no_269 <- full_50 %>% filter(sku_id == 269) %>% 
  select(placed_at, qty) %>% 
  rename(ds = placed_at, y = qty)%>% 
  group_by(ds) %>% 
  mutate(y = sum(y)) %>%
  unique() 

# Plot of quantity
no_269 %>% ggplot(aes(x = ds, y = y)) +
  geom_line()
```

### Prepping objects for time series modeling.
```{r}
# Trying to utilize the forecast package to forecast daily data
# The forecast package was not built to forecast daily data.

# Finding the date-time conversions for the start day and week
startW <- as.numeric(strftime(head(no_269$ds, 1), format = "%W"))
startD <- as.numeric(strftime(head(no_269$ds, 1) + 1, format =" %w"))

# Checking to see if the time series will work with the date transformation
print(ts(data = no_269$y, frequency = 7, start = c(startW, startD)), calendar = T)
```

```{r}
# Making a time series object
y <- ts(data = no_269$y, frequency = 7, start = c(startW, startD))

autoplot(y) + ggtitle("Time Plot: Quantity of SKU per Week") + ylab("Total of SKU Sold")

# Stationary transform the data
y2 <- diff(y)

# Viewing the transformed data: change in real data over time
autoplot(y2)

# Series appears trend-stationary (trend plot is flat)

# Looking for weekly trends
ggseasonplot(y2) + ylim(-50, 50)
```

### The Question: How can we model these data?

Three types of model algorithms offered in the `forecast` package are a seasonal naive model, exponential smoothing, and ARIMA models.
  - Seasonal naive (Predicts the same value as the previous time period)
  - Exponential smoothing ($y_{T+1}|T = \alpha y_T + \alpha (1 - \alpha) y_{T-1} +\alpha (1 - \alpha)^2 y_{T-2}+ ...$)
  - ARIMA (Autoregressive Integrated Moving Average)

In this markdown we try to forecast using these models in the `forecast` package on an SKU from our cleaned data.

### Exponential Smoothing
```{r}
### fitting the time series with exponential smoothing
fit_ets <- ets(y)

print(summary(fit_ets))

autoplot(fit_ets)

checkresiduals(fit_ets)

### Using the difference data

#fit_ets2 <- ets(y2)

#print(summary(fit_ets2))

#autoplot(fit_ets2)

#checkresiduals(fit_ets2)
```
It appears the residuals for these data are strictly positive and hence have an exponential distribution (see `checkresiduals(fit_ets)`). This is not ideal, and means this model is missing out on using some information.

### ARIMA
```{r}
fit_arima <- auto.arima(y, stepwise = FALSE, approximation = FALSE, trace = TRUE)

print(summary(fit_arima))

checkresiduals(fit_arima)
```

This model is not a great model, considering the p-value is 1, even though it was the 'best model' that the `auto.arima()` function chose.

### Comparing the two models.
```{r}
# Note that x axis has the week of the year, and anything past 62 is a new year (starting in January)
fcst_ets <- forecast(fit_ets, h = 56) # 56 days = 8 weeks

fcst_arima <- forecast(fit_arima, h =56)

# Feel free to comment out the ylims, I made them to better visualize the data, but it removed some of the outliers.
autoplot(fcst_ets) + ylim(0, 1000)

autoplot(fcst_arima)+ ylim(0, 500) 
```

