---
title: "Model Comparison"
author: "Carson Slater"
date: '2022-10-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r, include=TRUE}
# Loading packages
library(tidymodels)
library(tinytex)
library(stringr)
library(janitor)
library(glmnet)
library(lubridate)
library(knitr)
library(mosaic)
library(prophet)
```

```{r, eval=TRUE}
load("Data_Iteration2.Rdata")
```

# The Big Idea for this .Rmd

In my project, I have a forecasting model workflow with `Prophet`. The goal of this `.Rmd` is to create benchmark models and see how much better we can get Prophet to perform.

# Prepping the Data for the Modeling Process
```{r}
split <- by_day_filter %>% 
  rename(ds = placed_at, y = qty) %>% 
  group_by(sku_id) %>% 
  group_split()
```

# Additionally We Will Scale the Data [0,1]

I found that SKU's #757 (`split[[253]]`), #983 (`split[[293]]`), #1487 (`split[[356]]`), have either 338 or 339 observations, so these SKU's will be good to compare the time series modeling process on.

# Visualization for SKU's

We want to get a glimpse of what we are looking at. Hence, we know that we need to visualize the data for each of these goods.

```{r}
# 757
by_day_filter %>% 
  filter(sku_id == 757) %>% 
  ggplot(aes(x = placed_at, y = qty)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Date",
       y = "Volume Purchased") +
  theme_minimal()

# 983
by_day_filter %>% 
  filter(sku_id == 983) %>% 
  ggplot(aes(x = placed_at, y = qty)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Date",
       y = "Volume Purchased") +
  theme_minimal()

# 1488
by_day_filter %>% 
  filter(sku_id == 1488) %>% 
  ggplot(aes(x = placed_at, y = qty)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Date",
       y = "Volume Purchased") +
  theme_minimal()
```

`sku_id` 757 appears to be a good where most of the time, the total volume distributed per day is below 10. This means it might not be the most urgent good to forecase. The other two, `sku_id` 983 and 1488 appear to be more seasonal and unpredictable with trend changes.

# Building a Benchmark `Prophet` Model
```{r}
# This is going to be a model for SKU_id 1488.

# Preprocessing
df <-  split[[293]] 

# Creating an evaluation data frame
eval <- df %>% 
  filter(ds >= max(df$ds) - weeks(8)) %>% 
  complete(ds = seq.Date(min(ds), max(ds), by="day")) %>% 
  mutate(y = ifelse(is.na(y), 0, y))
  
eval <- eval %>% slice(1:(n() - 1))

orig_df <- df

df <- df %>% filter(ds < max(df$ds) - weeks(8))
```

```{r}
mod <- prophet(df, 
               yearly.seasonality = TRUE, 
               fit = FALSE, 
               interval.width = 0.8) %>% 
  add_country_holidays(country_name = 'ID')

fit <- fit.prophet(mod, df)

# 2 Months
future <- make_future_dataframe(fit, periods = 8*7)

# Make a forecast
fcst <- predict(fit, future)

##########################################################
# Since we cannot have negative predictive quantities, 
# I invert the upper confidence interval and set the 
# negative yhats to zero
##########################################################

fcst <- fcst %>% mutate(diff = yhat_upper - yhat)

fcst <- fcst %>%
  mutate(yhat_upper = ifelse(yhat_upper >= 0, yhat_upper, diff)) %>% 
  mutate(yhat = ifelse(yhat >= 0, yhat, 0), 
         yhat_lower = ifelse(yhat_lower >= 0, yhat_lower, 0))

prophet_plot_components(fit, fcst)

# Final plot of forecast
plot(fit, fcst) + 
  add_changepoints_to_plot(m = fit, cp_color = "orange") +
  labs(title = "Fitted Model and Forecast of an SKU",
       x = "Date",
       y = "Volume Purchased") +
  theme_minimal()
```

```{r}
# Model Evaluation
pred <- fcst %>% filter(ds >= max(orig_df$ds) - weeks(8)) %>%
  select(ds, yhat_lower, yhat, yhat_upper)

# Make a tibble of errors from the days
err_df <- tibble(pred$ds, eval$y, pred$yhat)

colnames(err_df) <- c("ds", "y", "yhat")

# creating a residual column
err_df <- err_df %>% mutate(resid = y - yhat)

# We need to filter zeros for the err_df so that we can find MAPE
# If the observed value is zero, MAPE is infinity.
err_df <- err_df %>% filter(y != 0)

#library(Metrics)

# Get RMSE
Metrics::rmse(err_df$y, err_df$yhat)

# RMSE = 11.8242

# Get SMAPE
Metrics::smape(err_df$y, err_df$yhat) # can be infinity if predicted values in denominator are close to 1

# SMAPE = .4353

# A MAPE function
mean(abs((err_df$y-err_df$yhat)/err_df$y))

# MAPE = .4366 - not super good.

err_df %>% 
  ggplot(aes(x = resid)) +
  geom_histogram(binwidth = 5, 
                 fill = "orange", 
                 color = "grey") +
  labs(title = "Residual Plot",
       x = "Residual Value",
       y = "Frequency") +
  theme_minimal()
```

### Attempting to Tune Parameters of `Prophet` Model

```{r}
# Plotting the original time series
#install.packages("TSstudio")
library(TSstudio)
library(xts)

# Plotting time series in base R plotting
ts <- xts(df$y, df$ds)

ts.plot(ts)

# What we are trying to predict (with ggplot2):
df %>% ggplot(aes(x = ds, y = y)) +
  geom_line() +
  xlab("") +
  scale_x_date(date_labels = "%Y %b %d")
```

# Testing Fourier Order
### Yearly Seasonality
```{r}
# This code chunk is simply to explore what different parameters in Prophet do.
# This is also going to be a model for sku_id 1488.

# This is not the actualy time series, but the yearly trend
test1 <- prophet(df, yearly.seasonality = 3)
prophet:::plot_yearly(test1)

test2 <- prophet(df, yearly.seasonality = 10)
prophet:::plot_yearly(test2)

test3 <- prophet(df, yearly.seasonality = 20)
prophet:::plot_yearly(test3)
```

### SECOND `Prophet` Model

In this I added changepoints into the model, so that it can know that the trend is switching. AS a result the model performed slightly better.

```{r}
# See plot for SKU #983 to see how I selected changepoints
mod2 <- prophet(growth = "linear",
                changepoints = c("2022-03-11", "2022-06-08"),
                yearly.seasonality = FALSE,
                daily.seasonality = FALSE,
                seasonality.prior.scale = 10,
                holiday.prior.scale = 10,
                changepoint.prior.scale = 10,
                mcmc.samples = 0,
                interval.width = 0.8,
                uncertainty.samples = 1000,
                fit = FALSE
                )

# Fitting the Model
fit2 <- fit.prophet(mod2, df)

# 2 Months
future <- make_future_dataframe(fit2, periods = 8*7)

# Make a forecast
fcst2 <- predict(fit2, future)

##########################################################
# Since we cannot have negative predictive quantities, 
# I invert the upper confidence interval and set the 
# negative yhats to zero
##########################################################

fcst2 <- fcst2 %>% mutate(diff = yhat_upper - yhat)

fcst2 <- fcst2 %>%
  mutate(yhat_upper = ifelse(yhat_upper >= 0, yhat_upper, diff)) %>% 
  mutate(yhat = ifelse(yhat >= 0, yhat, 0), 
         yhat_lower = ifelse(yhat_lower >= 0, yhat_lower, 0))

prophet_plot_components(fit2, fcst2)

# Final plot of forecast
plot(fit2, fcst2) + 
  add_changepoints_to_plot(m = fit, cp_color = "orange") +
  labs(title = "Fitted Model and Forecast of an SKU",
       x = "Date",
       y = "Volume Purchased") +
  theme_minimal()
```

```{r}
# Model Evaluation
pred2 <- fcst2 %>% filter(ds >= max(orig_df$ds) - weeks(8)) %>%
  select(ds, yhat_lower, yhat, yhat_upper)

# Make a tibble of errors from the days
err_df2 <- tibble(pred2$ds, eval$y, pred2$yhat)

colnames(err_df2) <- c("ds", "y", "yhat")

# creating a residual column
err_df2 <- err_df2 %>% mutate(resid = y - yhat)

# We need to filter zeros for the err_df so that we can find MAPE
# If the observed value is zero, MAPE is infinity.
err_df2 <- err_df2 %>% filter(y != 0)

#library(Metrics)

# Get RMSE
Metrics::rmse(err_df2$y, err_df2$yhat)

# RMSE = 7.603284

# Get SMAPE
Metrics::smape(err_df2$y, err_df2$yhat) # can be infinity if predicted values in denominator are close to 1

# SMAPE = .2466916

# A MAPE function
mean(abs((err_df2$y-err_df2$yhat)/err_df2$y))

# MAPE = .340391

err_df2 %>% 
  ggplot(aes(x = resid)) +
  geom_histogram(binwidth = 3, 
                 fill = "orange", 
                 color = "grey") +
  labs(title = "Residual Plot",
       x = "Residual Value",
       y = "Frequency") +
  theme_minimal()
```

After reviewing the model evaluation metrics, this model was slightly better than the prior one. I need to keep tuning parameters to get the model to have the least amount of error.